---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---



```{r}
library(dplyr)

names(gapminder)

subset_gapminder<-gapminder%>%select(year, country, gdpPercap)


dim(subset_gapminder)


year_country_gdp_euro<-gapmminder%>%
  filter(continent="Europe")%>%
select (year, country, gdpPercap)
year_country_gdp_euro




Africa_lifeexp<-gapminder%>%
  filter(continent=="Africa")%>%
  filter(year=="2007")

select (year, country, lifeExp)
Africa_lifeexp


head(gapminder)

str(gapminder)

gap_by_continent<-gapminder%>%group_by(continent)

gdp_by_continent<-gap_by_continent%>%
  summarize (mean_gdpPercap=mean(gdpPercap))

gdp_by_continent


lifeexp_by_africa_country<-gapminder%>%
  filter (continent=="Africa")%>%
  group_by (year) %>%
  summarize (mean_lifeexp_africa=mean(lifeExp))

lifeexp_by_africa_country

rmarkdown::render("diplyerandtidyRnotebook.Rmd","html_notebook")


gdp_by_continent<-gapminder%>%
  group_by(continent, year)%>%
  summarize(mean_gdp=mean(gdpPercap),
sd_gdp = sd(gdpPercap),
mean_pop = mean(pop),
sample_size = n(),
se_gdp = sd_gdp/sqrt(sample_size))

mean_pop


#combining dplyr with ggplot2#

gapminder%>%filter(continent=="Asia")%>%
  ggplot(aes(x=gdpPercap, y = lifeExp)) +
geom_point()
  

````{r}

gap_wide<-read.csv("data/gapminder_wide.csv", stringsAsFactors = FALSE)
colnames(gap_wide)


#This data is in a wide format, we want to change it to long#

head(gap_wide)
  

gap_long<-gap_wide%>%


gap_long<-gap_long%>% separate(obstype_year, into = c("obs_type", "year"), sep = "_"



gap_long$year<-as.integer(gap_long$year)
head(gap_long)

lifespan<-gap_long%>%  
group_by(continent)%>%
  summarize(meanlifespan= mean(lifeExp))
lifespan

gap_long%>% #filter(obs_type=="lifeExp")%>%
  group_by (continent)%>%
  summarize(mean_lifeExp = mean(obs_values))


gap_normal<-gap_long%>% spread (obs_type, obs_values)
str (gap_normal)
colnames (gap_normal)

names (gap)
gap_normal<-gap_normal[,names(gap)]


all.equal (gapminder, gap_normal)












gap_normal<-gap_normal%>%arrange(country, continent, year)

















``````







```

gap_long<-gap_wide%>%
gather (obstype_year, obs_values, -continent, -country)

















###working with data the dplyr way

select() - subset variables
filter() - subset observations (rows)
group_by() - counts and group data
mutate() - adda new column based on the data
group_by() function subset based on some column







